---
title: "Functions in R"
author: "Peter Nandori"
date: "2/8/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Writing your own function

Many functions are already available in R as default and some are available by installing special packages.
Still, you may want to write your own function for efficient execution of tasks. To this end, you need the
R function *function*. The following function takes $x,a,b$ as an input and compute $\log(a*x+b)$ as an output.

```{r}
##################################################################################
## toyFunc1
##   multiplying by a, adding 1 and then taking logarithm 
##  Input:
##   x: vector
##   a: some number 
##   b: some number
##  Output:
##   y: log(a x + b)
##################################################################################
toyFunc1 = function(x,a,b){
    y = log(a*x+b)
    return(y)
}
```

Here, 
**x, a, b** 
within the parentheses are inputs to the function named 
**toyFunc1**. 
After the parentheses, R
statements you want to execute are included in the curly brackets. 
For an output, the function return is used. If
you do not have a quantity to return, you can simply end statements without return.
Now we can invoke our function 
**toyfunction** for a list of $x$'s
with $a = 2$, $b = 100$. Note the
two different syntax
```{r}
x = 1:5
toyFunc1(x=x,a=2,b=100)
toyFunc1(x,2,100)
```

If you use $b=100$ almost always, it is desirable to set an default value:
```{r}
toyFunc2 = function(x,a,b = 100){
    y = log(a*x+b)
    return(y)
}
```

Since the default value is set,
we can now type
```{r}
toyFunc2(1:5,2)
```

Do you remember the previous homework
problem, where you needed to run the 
same code with the only difference
that a parameter $n$ was updated?

This should be easily done with a function.


```{r}
################################################################################
## simExp
##    Plots the histogram over N experiments.
##    Each experiment consists of 
##    standardizing the sample mean 
##    of a sample of size n from an exponential population.
##  Input:
##    N: number of experiments
##    n: sample size
##    rate: parameter of exponential distribution
##    
##  Output:
##    Histogram and plot of bell curve
#################################################################################
simExp = function(N, n, rate=2){
    xbars = numeric(N) 
    for(i in 1:N){
        dataSet =
          rexp(n=n,rate=rate) 
        xbars[i] = mean(dataSet) 
    }
    zs = (xbars-1/rate)*(rate*sqrt(n)) 
    hist(zs,prob=TRUE,
         ylim=c(0,.6),
         xlim=c(-4,4),
         main="Exponential Distribution",
         xlab=paste("n=",n)) 
    curve(dnorm(x,mean=0,sd=1), add=TRUE)
}

```

Now we can easily produce four 
plots, each with the function
**simExp** but with different $n$
```{r}
par(mfrow=c(2,2))
simExp(N=1000,n=5)
simExp(N=1000,n=10)
simExp(N=1000,n=20)
simExp(N=1000,n=40)
```


## Exercise 

Write a function **VarianceCI**
that does the following.

(1)
takes the number of simulations **N**, the sample size **n**, the parameters **mean** and **sdev**, level of confidence **alpha**
as inputs, 

(2) generates random samples of size 
**n** from the normal distribution with parameters **mean** and **sdev**, 

(3) compute
$(100-alpha)\%$ CI 
for the variance **sdev**^2
(you may want to use **qchisq**,
the quantile function of the 
chi-squared distribution),

(4) repeat (2)-(3) **N** times, 

(5) compute the successful coverage proportion, that is the 
number of times the CI actually
contains **sdev**^2, divided by **N**

(6) draw a plot for coverage using matplot as we did in the file
TDistrCofidenceIntervals.R last week.

Finally, run
this function with **mean** = 0 and **sdev** = 1
and with both
**N** =10, **N** =100,
for both **n** = 5, **n** = 100
(that is, 4 cases in total).
Discuss your
finding.

Submit your code as well as the resulting plots and discussions in an
R Markdown file.