---
title: "Chi-squared test"
author: "Peter Nandori"
date: "3/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chi-squared test in R

## Problem
The response time of a computer system to a request for a
certain type of information is hypothesized to have an
exponential distribution with parameter $1$ sec (so if $X =$ response time, the pdf of X
under $H_0$ is $e^{-x}$ for $x>0$)

### (a)
If you had observed and wanted to use the
chi-squared test with five class intervals having equal
probability under $H_0$, what would be the resulting class
intervals?
Hint: use *qexp(1/5)*, etc.

### (b)
Carry out the chi-squared test using the following data
resulting from a random sample of 40 response times:

```{r}
vec<-c(.10,.99,1.14,1.26, 3.24,.12,.26,.80,.79,1.16,
1.76,.41,.59,.27,2.22,.66,.71, 2.21,.68, .43,   
.11,.46,.69,.38,.91, .55,.81,2.51,2.77,.16,
1.11,.02,.13,.19,1.21,1.13,2.93,2.14,.34,.44) 
```

Hint: Construct a vector ${\bf count}$ so that ${\bf count[i]}$ contains the number of sample
elements that belong to the $i$th class interval found in part (a) for $i=1,...,5$. Then perform the chi-squared test 
using ${\bf chisq.test(count) }$. As always, you can type ${\bf help(chisq.test) }$ in RStudio
to learn more about the chi-squared test in R. 


# Solution

The following list contains the endpoints of the intervals

```{r}
quantiles <- c(qexp(0),qexp(1/5),qexp(2/5),qexp(3/5),qexp(4/5), qexp(1))
```

One possible solution is 

```{r}
bins = 5
count = rep(0,5)
for(i in 1: length(vec)){
  count[ceiling(bins*pexp(vec[i]))] = count[ceiling(bins*pexp(vec[i]))] +1
}
chisq.test(count)
```

















Let $X$ be an $m \times n$ data matrix such that $X^{T} X$ is invertible, and let $M=I_{m}-X\left(X^{T} X\right)^{-1} X^{T} .$ Add a column $\mathbf{x}_{0}$ to the data and form
$W=\left[\begin{array}{ll}{X} & {\mathbf{x}_{0}}\end{array}\right]$
Compute $W^{T} W .$ The $(1,1)$ -entry is $X^{T} X .$ Show that the Schur complement (Exercise 15$)$ of $X^{T} X$ can be written in the form $\mathbf{x}_{0}^{T} M \mathbf{x}_{0} .$ It can be shown that the quantity $\left(\mathbf{x}_{0}^{T} M \mathbf{x}_{0}\right)^{-1}$ is the $(2,2)$ -entry in $\left(W^{T} W\right)^{-1} .$ This entry has a useful statistical interpretation, under appropriate hypotheses.



