---
title: "T test"
author: "Peter Nandori"
date: "3/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## T test in R

For the Z-test, we can compute the type II error probability analytically.
However for the
t-test, the type II error probability cannot be given in a closed form because the
test statistic contains the sample standard deviation $S$.
What we can do, is to approximate
the type II error probability by simulations.

**General principle**: if we do not know
the probability $p$
of a random event $E$, 
then we can generate a lot of independent
experiments and count the relative 
frequency of $E$. This relative frequency
will be our estimate of $p$. This
strategy is called *Monte Carlo simulation* and is very common in
computational probability.



```{r}
##################################################################################
## tTestSim
##   Monte Carlo simulation of type I and type II error probabilities of T test
##  Input:
##   N: number of iterations
##   n: sample size 
##   mean: population mean
##   mean0: mu_0 in the null hypothesis
##   sigma: population standard deviation
##   test: type of alternate hypothesis
##  Output:
##   y: Monte Carlo simulation of type I / type II error probability
##################################################################################
tTestSim<-function(N=1000,n=20,mean=0,
        mean0=1,sigma=1,alpha=.05,seedNum=1,test="lowerTail"){
	## determine seed for reproducing a result for simulation
	set.seed(seedNum)
	## result vector
	rejections <- numeric(N)

	for(i in 1:N){
		## generate data from a normal distribution
		dataSet <- rnorm(n=n,mean=mean,sd=sigma)
		## compute test statistic
		xbar <- mean(dataSet)
		s <- sd(dataSet)
		testStat <- (xbar-mean0)/(s/sqrt(n))

		if(test=="lowerTail"){
			## H0: mu=mu0 vs Ha: mu < mu0
			## RR testStat < -t_{alpha,n-1}
			if(testStat < -qt(p=alpha,df=n-1,lower.tail=FALSE)){
				rejections[i] <- 1
			}
		}
	}

	## Reporting Results
	if(test=="lowerTail"){
		cat("Hypothesis Testing: H0: mu=",mean0," vs. Ha: mu <",mean0,"\n")
		if(mean>=mean0){
			cat("Simulation: Type I Error Probability = ",mean(rejections),"\n")
		}else{
			cat("Simulation: Type II Error Probability",1-mean(rejections)," at mu = ",mean,"\n")
		}
	}
}
```


Let us now test this function:

```{r}
tTestSim()

```


```{r}
tTestSim(mean = 0.9)
```

As you can see, the function is incomplete
as only the lower tail part is implemented. Please implement the other
$2$ cases as a homework.



```{r}
##################################################################################
## tTestSim
##   Monte Carlo simulation of type I and type II error probabilities of T test
##  Input:
##   N: number of iterations
##   n: sample size 
##   mean: population mean
##   mean0: mu_0 in the null hypothesis
##   sigma: population standard deviation
##   test: type of alternate hypothesis
##  Output:
##   y: Monte Carlo simulation of type I / type II error probability
##################################################################################
tTestSim<-function(N=1000,n=20,mean=0,
        mean0=1,sigma=1,alpha=.05,seedNum=1,test="lowerTail"){
	## determine seed for reproducing a result for simulation
	set.seed(seedNum)
	## result vector
	rejections <- numeric(N)

	for(i in 1:N){
		## generate data from a normal distribution
		dataSet <- rnorm(n=n,mean=mean,sd=sigma)
		## compute test statistic
		xbar <- mean(dataSet)
		s <- sd(dataSet)
		testStat <- (xbar-mean0)/(s/sqrt(n))

		if(test=="lowerTail"){
			## H0: mu=mu0 vs Ha: mu < mu0
			## RR testStat < -t_{alpha,n-1}
			if(testStat < -qt(p=alpha,df=n-1,lower.tail=FALSE)){
				rejections[i] <- 1
			}
		}else if(test=="upperTail"){
		  ## H0: mu=mu0 vs Ha: mu > mu0
			## RR testStat > t_{alpha,n-1}
		  if(testStat > qt(p=alpha,df=n-1,lower.tail=FALSE)){
				rejections[i] <- 1
			}
		}else if(test=="twoSided"){
		  ## H0: mu=mu0 vs Ha: mu =/= mu0
			## RR |testStat| >= t_{alpha/2,n-1}
		  if(abs(testStat) >= qt(p=alpha/2,df=n-1,lower.tail=FALSE)){
				rejections[i] <- 1
			}
		}
	}

	## Reporting Results
	if(test=="lowerTail"){
		cat("Hypothesis Testing: H0: mu=",mean0," vs. Ha: mu <",mean0,"\n")
		if(mean>=mean0){
			cat("Simulation: Type I Error Probability = ",mean(rejections),"\n")
		}else{
			cat("Simulation: Type II Error Probability",1-mean(rejections)," at mu = ",mean,"\n")
		}
	}else if(test=="upperTail"){
	  cat("Hypothesis Testing: H0: mu=",mean0," vs. Ha: mu >",mean0,"\n")
	  if(mean<=mean0){
			cat("Simulation: Type I Error Probability = ",mean(rejections),"\n")
		}else{
			cat("Simulation: Type II Error Probability",mean(rejections)," at mu = ",mean,"\n")
		}
	}else if(test=="twoSided"){
	  cat("Hypothesis Testing: H0: mu=",mean0," vs. Ha: mu =/=",mean0,"\n")
	  if(mean==mean0){
			cat("Simulation: Type I Error Probability = ",mean(rejections),"\n")
		}else{
			cat("Simulation: Type II Error Probability",mean(rejections)," at mu = ",mean,"\n")
		}
	}
}
```


```{r}
tTestSim(mean = 0.9,test="lowerTail")
```

```{r}
tTestSim(mean = 0.9,test="upperTail")
```

```{r}
tTestSim(mean = 0.9,test="twoSided")
```

```{r}
tTestSim(mean = 1.1,test="lowerTail")
```

```{r}
tTestSim(mean = 1.1,test="upperTail")
```

```{r}
tTestSim(mean = 1.1,test="twoSided")
```



